$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-04-30 paper/paper.tex
Small fixes here and there

\subsubsection{Typing Contexts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{nidefinition}
  Figure \ref{fig:prectx} defines types as either unit types, base types or channel types.
  The unit type $\unit$ serves as a proof of inhabitance for types (we can trivially obtain it), and is used to such end in subject congruence --- for the $\new \PO_{\suc n} \eqeq \PO_n$ rule.
  The base type $\base{n}$ uses a natural number $n$ to represent types.
  This natural number can then be interpreted as a type in the host language.
  Using natural numbers instead of type universes is convenient: less universe polymorphism poluting specifications.

@@

\subsubsection{Usage Contexts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We choose to allow multiple multiplicity algebras to coexist in our typing system.
We do so by defining an indexed set of usage algebras.

\begin{nidefinition}
  An indexed set of usage algebras is a set of indices such that for every index there exists a carrier $C$ with a $\Quantifier_C$ algebra defined on it, as captured by Figure \ref{fig:indexed-multiplicities}.
  We require the index type to be inhabited -- proving subject congruence on $\new \PO_{\suc n} \eqeq \PO_n$ requires an inhabited carrier set.

@@

We make use \emph{leftover typing} \cite{}, an approach that, in addition to the usual typing context and input usage context, adds an extra output usage context to the typing rules.
This output context contains the \emph{leftovers} of the process being typed --- its unused multiplicities.
Crucially, the leftovers can then be used as an input to some othew typing derivation.
Adopting this techinque has two main advantages:
\begin{itemize}
  \item
    The input and output actions do no longer need to transform the usage context through functions, but can use a variable reference type with both an input and an output usage context to encapsulate these transformations.
  \item
    Parallel composition no longer needs an explicit context splitting proof --- a proof $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$ where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
    Such a proof could be obtained through automated search --- but it is better yet to avoid needing it.
    With leftover typing, we use the leftovers of $P$ to type $Q$.
\end{itemize}

>>>>

\begin{nidefinition}
  We define types (Figure \ref{fig:prectx}) as either unit types, base types or channel types.
  The unit type $\unit$ serves as a proof of inhabitance for types (we can trivially obtain it), and is used to such end in subject congruence --- for the $\new \PO_{\suc n} \eqeq \PO_n$ rule.
  The base type $\base{n}$ uses a natural number $n$ to represent types.
  This natural number can then be interpreted as a type in the host language.
  Using natural numbers instead of type universes is convenient: less universe polymorphism poluting specifications.

@@

\begin{nidefinition}
  We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
  An indexed set of usage algebras is a set of indices such that for every index there exists a carrier $C$ with a $\Quantifier_C$ algebra defined on it, as captured by Figure \ref{fig:indexed-multiplicities}.
  We require the index type to be inhabited -- proving subject congruence on $\new \PO_{\suc n} \eqeq \PO_n$ requires an inhabited carrier set.

@@

We make use \emph{leftover typing} \cite{}, an approach that, in addition to the usual typing context and input usage context, adds an extra output usage context to the typing rules.
This output context contains the \emph{leftovers} of the process being typed --- its unused multiplicities.
Crucially, the leftovers can then be used as an input to some other typing derivation.
Adopting this techinque has three main advantages:
\begin{itemize}
  \item
    The input and output actions no longer transform the usage context through functions.
    Instead, they use a variable reference type with both an input and an output usage context that encapsulates these transformations.
  \item
    Parallel composition no longer needs an explicit context splitting proof --- a proof $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$ where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
    Such a proof could be obtained through automated search --- but it is better yet to avoid needing it.
    With leftover typing, we use the leftovers of $P$ to type $Q$.
  \item
    Theorems like framing (\S \ref{framing}), weakening (\S \ref{weakening}) and strengthening (\S \ref{strengthening}) become suddenly stateable.
\end{itemize}
